A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2015-11-24 docs/discussion_v3.02.tex
Started rewriting the discussion.

We compare our findings of the stellar mass-to-light ratios with the study by \cite{SWELLSV}: They found that the bulge of J1331 has an IMF \emph{more} bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8''$. In Section \ref{sec:results_JAM_NFW} we fitted a mass model with NFW halo to (1) the Einstein mass and (2) stellar kinematics within $\sim 3.5''$. Our best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a \emph{less} bottom-heavy IMF than the Salpeter IMF. \cite{SWELLSV} found systematically lower NFW halo masses ($v_\text{circ,halo}(5'') \sim 120~\text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ,halo}(5'') \sim 200~\text{km s}^{-1}$).

@@

In Section \ref{sec:results_JAM_SB} we created a dynamical model from only the surface brightness distribution and an increasing mass-to-light ratio profile without additional DM halo. We found that such a model would be perfectly consistent with the lensing results and predict a total $\Upsilon_\text{I,tot}(R'\sim0) = 2.53$ in the center of J1331. This is very close to the stellar $\Upsilon_\text{I,*}^\text{Chabrier} = 2.5 \pm 0.6$ based on the estimates by \citet{SWELLSI} for the bulge stellar mass using the Chabrier IMF , i.e., less bottom-heavy (more bluish) population.

@@

The above discussion motivates the following speculation: In the absence of a strong DM contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy IMF close to the Salpeter IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a central stellar population with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results. (According to Figure \ref{fig:lenscompareboth} the lensing result might not predict a mass to light ratio gradient inside $R_\text{ein}$, but then again the mass slope $alpha=1$ was only weakly constrained.) The disk of J1331 has a lower $\Upsilon_\text{I,*}$ than the bulge (see Table \ref{tab:previousresults} according to \citet{SWELLSI}). Such a drop in $\Upsilon_\text{I,*}$ at the transition region from bulge to disk around $\sim 5''$ could lead to the observed drop in $v_\text{rms}$, while an increasing contribution of a lower mass DM halo at larger radii as found by \citet{SWELLSV}, would recover the kinematics in the out regions of J1331's disk.

>>>>

We compare our results from Section \ref{sec:results_JAM_NFW} with the study by \citet{SWELLSV}: They found that the bulge of J1331 has an IMF \emph{more} bottom-heavy than the Salpeter IMF from fitting a NFW halo to (1) the Einstein mass and (2) gas kinematics at larger radii $\gtrsim 8''$. In Section we fitted a mass model with NFW halo to (1) the Einstein mass and (2) stellar kinematics within $\sim 3.5''$. Our best fit $\Upsilon_\text{I,*} = 4.2 \pm 0.2$ (see Table \ref{tab:modelB4_bestfit}) indicates a \emph{less} bottom-heavy IMF than the Salpeter IMF. \cite{SWELLSV} found systematically lower NFW halo masses ($v_\text{circ,halo}(5'') \sim 120~\text{km s}^{-1}$ according to their Figure 2) than we did ($v_\text{circ,halo}(5'') \sim 200~\text{km s}^{-1}$).

We compare our result from Section \ref{sec:results_JAM_SB} with the findings from \citet{SWELLSI}: We created a dynamical model from only the surface brightness distribution and an increasing mass-to-light ratio profile without additional DM halo. We found that such a model would be perfectly consistent with the Einstein mass and predict a total $\Upsilon_\text{I,tot}(R'\sim0) = 2.53$ in the very center of J1331. This is close to the stellar $\Upsilon_\text{I,*}^\text{Chabrier} = 2.5 \pm 0.6$ based on the estimates by \citet{SWELLSI} for the bulge stellar mass using the Chabrier IMF , i.e., less bottom-heavy (more bluish) population.

@@

@@

\subsection{On J1331's possible merger history}

\Wilma{[TO DO: Remove minor in merger.]}

J1331 has a large counter-rotating stellar core within $\sim 2''$. This suggests a process in J1331's past in which two components with angular momenta oriented in opposite directions were involved. Accretion of gas on retrograde orbits and subsequent star formation could lead to a younger and counter-rotating stellar population \Wilma{[TO DO: REF]}. However, to form enough stars such that the net rotation of the large and massive core is retrograde, a very large amount of gas would have had to be accreted by J1331---which is not very likely [TO DO: REF].

Another scenario are galaxy mergers. Major mergers including large amounts of gas can form kinematically decoupled cores (KDCs), see e.g., \Wilma{[TO DO: REF: Lauren Hofman et al 2010 - read paper how this fits with the reverse color gradient]}. During a minor merger, the dense nucleus of a satellite galaxy on a retrograde orbit could survive the dissipationless accretion and could spiral to the core due to tidal friction \citep{1984ApJ...287..577K}. Usually large bulges of spirals or ellipticals are reddest in the center and become bluer further out.  Because star formation in low-mass satellite galaxies occurred later than in massive galaxies \Wilma{[TO DO: REF]}, the merger remnant would now host a younger and bluer population with lower velocity dispersion within the older and redder bulge of the massive progenitor. Such a reverse color gradient could correspond to an increasing stellar mass-to-light ratio.

The merger theory could explain and support our findings from dynamical modelling and the previous sections, that J1331 might have an increasing stellar mass-to-light ratio within its core.

\subsection{On possible modelling failures and speculations how to resolve them}

\Wilma{[TO DO: Include in discussion: Kinematic twist due to merger. Bulge and disk have different kinematic major axis, and the measured velocities in the disk, which are off from the major axis, are much lower than the actual maximum vcirc. Could explain the dips in the outer regions.]}

Another effect that minor mergers can have on galaxies are a misalignement (warp) of the photometric and kinematic major axis \Wilma{[TO DO: REF]}. Would this be the case for J1331 the assumption of axisymmetry is not true anymore and our dynamical modelling would fail.

Mergers also could have changed the 3D shape of the dark matter halo and the NFW halo was therefore not a good model for J1331's dark matter halo. We also tried to model J1331 with a cored logarithmic halo. However, due to degeneracies in the modelling, we were not able to either constrain the profile for a cored logarithmic halo, or to rule it out.

\Wilma{[TO DO: Include in this discussion the findings from the color profile.]}

\subsection{Future work}

Standard JAM modelling approaches seem not to work for J1331. A JAM model for J1331 would need to allow for stellar mass-to-light ratio gradients within the galaxy, velocity anisotropy and a dark matter halo. Because of degeneracies between stellar and dark mass, and matter distribution and anisotropy profile, such a dynamical model would not lead to very tight constraints on the model parameters.  A search for colour gradients in J1331 and/or investigation of absorption line indices could support or contradict the suspicion of the existence of stellar mass-to-light ratio gradients in J1331. Subsequently detailed stellar population analyses of the spectra taken along J1331's major axis should be conducted to constrain the mass-to-light ratio reliably.  \Wilma{[TO DO: Include in this discussion the findings from the color profile.]}

In addition, the dynamical modelling should use more of the available information on J1331 and fit dynamics (stellar and gas kinematics from \citet{SWELLSV}) simultaneously with the gravitational lensing (image positions, shape and even flux ratios) in a similar fashion to \citet{SWELLSIV}. To also model the extent, shape and flux of the lensing images, the method by \citet{2004ApJ...611..739T,2003ApJ...590..673W} could be employed, which models the surface brightness distribution of the images and source on a pixelated grid. However, for this to work a good model for the galactic extinction would be needed.

High-resolution integral-field spectroscopy could help with this, allow for spatially resolved stellar population analysis and dynamical modelling in two dimensions. \Wilma{[TO DO: Include in this discussion that this could also help test, if the vrms dip is maybe because of twist, i.e., the measurements where not done along the major axis in the disk]} This would lead to much better understanding of J1331's structure and mass distribution and therefore answer questions on how minor mergers might modify spiral galaxies.

\Wilma{[TO DO: Include in discussion: High resolution IFU: high spatial resolution --> to resolve small features. High spectral resolution --> because the velocity dispersion in the outer regions of the galaxy are very low and to measure them accurately, need high spectral resolution.]}

\Wilma{[TO DO: Include the following discssion somwhere: The above discussion motivates the following speculation: In the absence of a strong DM contribution in the center, the overall stellar-mass-to-light ratio within the Einstein radius indicates a bottom-heavy IMF close to the Salpeter IMF, consistent with estimates from the velocity dispersion. At the same time the central dip can then be only explained, if there was an increase in stellar mass-to-light ratio with radius \textit{within} the bulge. \emph{If} there was a central stellar population with an IMF close to the \citet{Chabrier2003} IMF, surrounded by a more bottom-heavy population, the central stellar kinematics would be well explained and be fully consistent with the lensing results. (According to Figure \ref{fig:lenscompareboth} the lensing result might not predict a mass to light ratio gradient inside $R_\text{ein}$, but then again the mass slope $alpha=1$ was only weakly constrained.) The disk of J1331 has a lower $\Upsilon_\text{I,*}$ than the bulge (see Table \ref{tab:previousresults} according to \citet{SWELLSI}). Such a drop in $\Upsilon_\text{I,*}$ at the transition region from bulge to disk around $\sim 5''$ could lead to the observed drop in $v_\text{rms}$, while an increasing contribution of a lower mass DM halo at larger radii as found by \citet{SWELLSV}, would recover the kinematics in the out regions of J1331's disk.]}
